---
title: "solar-plant-power-generation"
author: "Alex McIntosh"
date: "7/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # Default to false considering I hope to use mostly graphs
  message = FALSE,
  warning = FALSE,
  tidy = TRUE,
  cahe = TRUE, # Set to FALSE to conserve memory
  eval = FALSE # Set to TRUE for code to actually rune
)
```

# Introduction

For more information about how the graphs and results were generated for this
report, see <http://www.github.com/mcintalmo/solar-plant-power-generation> for
both the R Markdown that generated this report and the notebook that used for
exploration.

# Methods and Analysis


## Data Collection
The data for this analysis was collected by two solar plants in India.

The first plant, id: 4135001 is near Gandikotta, Andhra and the second plant,
id: 4136001 is near Nasik, Maharashtra.

Because of the color schemes of the graphs to follow, for the remainder of the 
report I will refer to the Gandikotta plant (4135001) as "Blue Plant" and plant
in Nasik (4136001) as "Red Plant."

The data were collected at 15 minute intervals for 34 days.

22 inverter sensors and one weather sensor at each plant brought the total to 46
senesors. 


The inverter sensors collected the DC and AC output of the group of solar panels
it was monitoring.

Additionally, the inverters tallied the daily yield in DC output from
midnight to midnight and maintained total yield the sensor had observed.

Total yield increased until a sensor was replaced, at which point it would begin
again at 0.


The weather sensors recorded the ambient temperature, the temperature of the
modulte, and the irradiation level.

Owing to the sensors' reading of 0 at night, it is assumed any irradiation is
owed to the sun's radiation.


THe data were 

## Data Cleaning



## Data Exploration
```{r pressure}
plot(pressure)
```


## Correlation


# Results


## Model Root Mean Squared Errors


## Validation


# Conclusion

